\section{Concluding Remarks \& Related Work}
\label{sec:relatedwork}
In this section we draw some conclusions based on the related work for correctness of code involving RCU memory management.

\mypar{Modeling RCU}
Alglave et al.~\cite{Alglave:2018:FSC:3173162.3177156} propose a memory model to be assumed by the platform-independent parts of the Linux kernel, regardless of the underlying hardware's memory model.
As part of this, they give the first formalization of what it means for an RCU implementation to be correct (previously this was difficult to state, as the guarantees in principle could vary by underlying CPU architecture). Essentially, that reader critical sections do not span grace periods.  They prove by hand that the Linux kernel RCU implementation~\cite{DBLP:conf/cav/AlglaveKT13,abssem} satisfies this property. 
According to their definition, our model in Section \ref{sec:semantics} is a valid RCU implementation (assuming sequential consistency).
To the best of our knowledge, ours is the first abstract \emph{operational} model for a Linux kernel-style RCU implementation --- others are implementation-specific~\cite{Mandrykin:2016:TDV:3001219.3001297} or axiomatic like Alglave et al.'s. Tassarotti et al. models a well-known way of implementing RCU synchronization without hurting readers' performance, \textsf{Quiescent State Based Reclamation}(QSBR)~\cite{urcu_ieee} where synchronization in between the writer thread and reader threads provided via per-thread counters. A direct implementation of our semantics would yield unacceptable performance, since both entering(\lstinline|ReadBegin|) and exiting(\lstinline|ReadEnd|) modify shared data structures for the \textit{bounding-threads} and \textit{readers} sets. A slight variation on our semantics would use a bounding set  that tracked such a snapshot of counts, and a vector of per-thread counts in place of the reader set. Blocking grace period completion until the snapshot was strictly older than all current reader counts would be more clearly equivalent to these implementations. Our current semantics are simpler than this alternative, while also equivalent. We also know that if we had this slight variation in our model then all actions(traverse and read a heap value) from a reader thread would \textit{reduce} to a single \textit{atomic} action: read of a data value. However, we do not prove the \textit{atomicity} of the readers actions within the scope of this paper.

\mypar{Program Logics} Fu et al.~\cite{shao_temp} extend Rely-Guarantee/Separation-Logic ~\cite{vafeiadis07,Feng:2007:RCS:1762174.1762193,Feng:2009:LRR:1480881.1480922} with the \textit{past-tense} temporal operator to eliminate the need for using a history variable and lift the standard separation conjuction to assert over on execution histories. Gotsman et al.~\cite{Gotsman:2013:VCM:2450268.2450289} take assertions from temporal logic to separation logic~\cite{vafeiadis07} to capture the essence of epoch-based memory reclamation algorithms and have a simpler proof than what Fu et al. have~\cite{shao_temp} for Micheal's non-blocking stack~\cite{Michael:2004:HPS:987524.987595} implementation under a sequentially consistent memory model.

\mypar{Batched-Deallocation} Gotsman et al.~\cite{Gotsman:2013:VCM:2450268.2450289} and Tassarotti et al.~\cite{verrcu} consider the existence of \textit{batched-deallocation}. There is an direct mapping between Gotsman et al.'s \textsf{detached} and \textsf{in\_sum} sets for hazard pointers model and our logical abstract \textit{free list}. In this paper we relate our logical \textit{free list} to the set of \textit{bounding-threads} in concrete RCU state to reason about the existence of the \textit{grace} period. In the case of using our system for hazard pointers, there would be a direct relation in between relation between our logical \textit{free-list} and the set of \textsf{detached} in Gotsman et al.~\cite{Gotsman:2013:VCM:2450268.2450289}. This leads us to think that it is a good future direction to investigate the applicability of our type system to different clients of different \textit{deferred} memory reclamation schemes~\cite{Wen:2018:IMR:3178487.3178488,Michael:2004:HPS:987524.987595,Mckenney:2004:EDD:1048173,UCAM-CL-TR-579}.
%
%This crucial mapping implies that our type system can also directly be applied to hazard pointers and we think that it is a good future direction to investigate the applicability of our type system to different clients of different \textit{deferred} memory reclamation schemes~\cite{Wen:2018:IMR:3178487.3178488,Michael:2004:HPS:987524.987595,Mckenney:2004:EDD:1048173,UCAM-CL-TR-579}.% Mandrykin et al.~\cite{Mandrykin:2016:TDV:3001219.3001297} use \textsf{VCC}~\cite{Cohen:2009:VPS:1616077.1616080} to verify an implementation of RCU primitives using specifications relating to a specific implementation, and do not verify client code.

\mypar{Weak Memory Models} Tassarotti et al.~\cite{verrcu} uses a protocol based program logic based on separation and ghost variables called \textsf{GPS}~\cite{Turon:2014:GNW:2660193.2660243} to verify user-level \textsf{RCU} singly linked list under \emph{release-acquire} semantics, which is a weaker memory model than sequential-consistency. They require \textit{release-writes} and \textit{acquire-reads} to the QSRB counters for proper synchronization in between the mutator and the reader threads. This synchronization is exactly the only one that we enforce in the logical observations of the mutator thread: from \textsf{unlinked} to \textsf{freeable}. Tassarotti et al.'s synchronization for linking a new node(Release-Acquire Pair 1) is also captured by our logical \textsf{fresh} observation. Based on these observations, we find it as an important future direction to investigate our type system formally with respect to these \textit{acquire-release} relations.
%They enforce the safe synchronization protocol in between the readers and the writer thread through \textit{release-writes} and \textit{acquire-reads} of the counters as a part of the QSRB model. We do capture all the guarantees from these \textit{release-acquire} orderings in our model via \textit{logical observations} -- e.g. (\textsf{iterator} $tid$, \textsf{unlinked} and \textsf{freeable}) -- the writer thread makes on the heap. Based on our comparative understanding taking Tassarotti's work as a reference, although we do not show it formally within the scope of this paper, we do not see any fundamental reason for not having our model to work under \textit{release-acquire} semantics.

\mypar{Separation of Concerns for Client Verification}: We think that reusability of predicates used in arbitrary client verification is important. In this regard, Tassarotti et al.'s \textit{abstract-predicates} -- e.g. WriterSafe -- might not be easy to reuse for an arbitrary RCU client -- e.g. binary search tree -- as they include linked list specific logical constructions together with the constructions for QSRB model. Unlike, we use similar types both for BST(Appendix \ref{appendix:bst_del}) and linked-list based bag example.
%We think that instantiating a basis logic requires an effort on providing clean separation between client verification and full functional correctness of a model. \textit{This effort is subtle and important}. Meyer et al. ~\cite{myr} presents a methodology to separate \textit{linearazibility} proofs for Michael\&Scott's queue~\cite{Michael:1996:SFP:248052.248106} and DGLM~\cite{Doherty:2004:DSB:1007912.1007945} from memory reclamation model via establishing an over approximation for the \textit{epoch-based} and \textit{hazard-pointers} reclamation model. We hope our generic abstractions for linked-data structures inspire them when they integrate linked-data structure abstractions(e.g. linked-list) to their methodology. Regarding specifying client's action, Tassarotti et al. provide \textit{abstract-predicates} -- e.g. WriterSafe -- which might not be so reusable for an arbitrary RCU client -- e.g. binary search tree -- as they include linked list specific logical constructions together with the constructions for QSRB model. On the other hand, we manage to keep our types for client specification not only relatively simple but also reusable for different clients (e.g. a linked list in Appendix \ref{appendix:bag_paul} and a binary search tree in Appendix \ref{appendix:bst_del} via careful instantiation of Views framework: logical state including thread's observations, global memory invariants in Appendix \ref{sec:lemmas} and the denotations of our types in Figure \ref{fig:denotingtypeenviroment} Section \ref{sec:soundness}.
%\mypar{Realizing Our Abstract RCU Model} Efficient implementations minimizes the cost of the read-side critical section operations: entering/exiting to the read-side critical section. Typically, implementations are structured so each reader thread (or CPU) maintains its own \emph{fragment} of the shared global state, which is only updated by that thread, but is read by the write-side primitives. For example, Tassarotti et al.~\cite{verrcu} model the readers' counters as an array of counters where the content of each cell corresponds to the counter of a specific reader thread in QSRB model. On the other hand, a direct implementation of our semantics would yield unacceptable performance, since both entering(\lstinline|ReadBegin|) and exiting(\lstinline|ReadEnd|) modify shared data structures for the \textit{bounding-threads} and \textit{readers} sets. A slight variation on our semantics would use a bounding set  that tracked such a snapshot of counts, and a vector of per-thread counts in place of the reader set. Blocking grace period completion until the snapshot was strictly older than all current reader counts would be more clearly equivalent to these implementations. Our current semantics are simpler than this alternative, while also equivalent.However, we know that if we had this slight variation in the implementation for the \lstinline|ReadBegin| and \lstinline|ReadEnd| then all actions(traverse and read the node value) from a reader thread would \textit{reduce} to an \textit{atomic} read of a data value(logical atomicity of readers' actions). Since we do not provide the careful implementation for the the \lstinline|ReadBegin| and \lstinline|ReadEnd| we cannot provide the \textit{atomicity} proof as well.
%(\textit{\textbf{RCU Model and Model Checking}}) - %Alglave et al. gives first semantics for the Linux kernel's implementation of \textsf{RCU}~\cite{abssem,Alglave:2018:FSC:3173162.3177156,DBLP:conf/cav/AlglaveKT13}. It captures details(behaviours on memory models etc.) of \textsf{RCU} implementation in Linux Kernel so it is much detailed and close to real implementation than our semantics is. However, our semantics abstracts the one Alglave et al. presents without loosing the \textit{fundamental law of} \textsf{RCU}  ~\cite{Alglave:2018:FSC:3173162.3177156} against which we show that our type system enforces correct usage of \textsf{RCU} API to guarantee memory safety(proper reclamation) and properties like \textit{acyclicity} for data structures implemented with using the API soundly.

%\mypar{Model Checking}
%Desnoyers et al. propose a virtual architecture to model out of order memory accesses and instruction scheduling~\cite{Desnoyers:2013:MSM:2506164.2506174} and use the \textsf{SPIN} model checker to verify a user-mode implementation of RCU~\cite{urcu_ieee}. This method requires manual translation from C to SPIN modelling language.
%Liang et al. presents an approach to verify the \emph{grace period} guarantee of \textsf{Tree RCU} in Linux kernel~\cite{LiangMKM16}. \textsf{Tree RCU} is an hieararchical \textsf{RCU} model which is a remedy to contention of writer threads in the single global lock model. 
%Similarly, Kokologiannakis et al.~\cite{Kokologiannakis:2017:SMC:3092282.3092287} use a stateless model checker for testing the core of \textsf{Tree RCU}.
%Liang's approach has limited support for lists and callback handling which are important when we consider the correctness for writer primitives which are based on callback handling. Kokologiannakis's approach overcomes all these limitations and shows considerable performance when compared with Liang's approach using \textsf{CBMC}. 
%Both focus on validating a particular RCU implementation, whereas we focus on verifying memory safety of clients independent of implementation.
 
%\mypar{Type Systems}
%Howard et al.~\cite{Howard:2011:RES:2001252.2001267} enhance \textsf{STM} in Haskell to a technique called \textit{Relativistic Programming} which aims low overhead linearly scalable concurrent threads. Unlike our system which types binary search tree in \textsf{RCU} setting, they claim  handling trees (look-up followed by update) as a future work.
%A follow up paper from Cooper et al.~\cite{Cooper2015RelativisticPI} presents a \textsf{Haskell} library called \emph{Monadic RP}, for relativistic programming. Similar to our approach, Monadic RP provides types and relativistic programming constructs for write/read critical sections which enforce correct usage of relativistic programming pattern. 
%They also have only checked a linked list.
%Thus our work is the first type system for ensuring correct use of RCU primitives that is known to handle more complex structures than linked lists.
